\documentclass{article}
\usepackage{times}
\usepackage[margin=1.25in]{geometry}
\renewcommand{\baselinestretch}{1.5} 

\usepackage[
    style=authoryear-icomp,
    natbib
]{biblatex}
\addbibresource{argumentative_singer.bib}

\usepackage[UKenglish]{babel}
\usepackage[autostyle]{csquotes}
\MakeOuterQuote{"}
\newcommand{\q}[1]{``#1''}

\usepackage{enumitem}
\setlist{parsep=0pt,listparindent=\parindent,}



\begin{document}
	\begin{flushright}
		University of Oslo
		\\\textbf{Course code:} EXPHIL03E
		\\\textbf{Candidate number:} 8025
		\\\textbf{Date:} 3rd December 2019
	\end{flushright}
\section*{Final submission of semester assignment}
Most human beings are speciesist, says Peter Singer (Singer 2 2017: 193). Discuss if Singer’s spesiecism argument includes not slaughtering animals and eating meat.%%ignorewc
		\begin{flushright}
			\textbf{Word count:} 932 words
		\end{flushright}
		\section*{Peter Singer, a defender of all sentient life}
		
		Peter Singer is an utilitarist philosopher who analyses a series of discussions which have been raised in the modern society, such as feminism topics, animal activist issues, and taking care of the environment, among others.

		\subsubsection*{Utilitiarism: definition}
		
		Utilitarism was defined previously by authors like Jermey Bentham and John Stuart Mill as the concept of choosing the path that would bring \emph{the greatest amount of good for the greatest number} \autocite{definitionutilitarism}. Nevertheless, this definition usually only considered the greatest number of \emph{people}, and not the greatest number of \emph{sentient beings}, as Peter Singer suggests. This was asked previosuly by Bentham, too, by questioning \q{who can feel pain}, claiming this is what matters.

		Considering sentient beings instead of only people, or even human beings\footnote{Excluding unconscious human benigs such as babies, fetuses, and possibly those in a coma. \autocite{singerpracticalethics}}, is a revolutionary thought, and has inspired a lot of vegetarians and vegans. After all, it is an anthropocentric thought to believe only human beings or people should have an ethical value. There is no rational reason which explains why humans should matter more than non-human beings from a consequentialist point of view. We can therefore take this as a (renewed) starting point for our ethical ideologies.

		\subsubsection*{Non-antrophocentric philosophies}
		
		In opposition to the previous ideas, we have pathocentric and biocentric views (ignoring the hollistic ideas), which, respectively, considers all sentient beings, and all living beings. Considering all sentient beings makes sense from a historical point of view, since not all human beings have been considered people throughout the years: Women and racialised people have not been considered subjects (as Kant puts it) since the beginning of time, but mere living beings that acts as means for men's purposes. This is still true in modern society for other animals, be it pets such as dogs and cats, or livestock animals, such as cows and pigs. 

		With this, we can draw a parallel between sexism, racism, and speciesism. If one of them is bad, the rest must by analogy also be bad, since there is no objective difference between one form of discrimination and another. \autocite{singerequal}

		Considering all biological life, nevertheless, is harder to argue for. For if not being able to feel pain, why would it matter how we treat them? From a purely deontological point of view, it would make sense to defend all forms of life, if our sense of what is good and correct were determined by \q{defending all forms of life}, but this is dogmatic from an utilitarian perspective. 

		On the other hand, since consequentalism tries to gain the highest benefit, it could be rational to defend all forms of life since this benefits us (all sentient beings, and their ecosystem) most.

		\subsubsection*{Types of utilitiarianism} 

		In accordance to \citet{smarttypes}, we can differ between two types of utilitarianism, and they each have their advantages and disadvantages. 

		One could say rule-utilitarianism is quite similar to Kant's deontology \autocite{kant}, but differs in one important aspect: the right action is done because it is right from a rational standpoint, as it gains the most pleasure for the most people in \emph{most} cases, and not because the action itself is inherently good.

		Combining them, one could intuitively understand that while killing an animal is wrong, per se, in most cases; if this means that it would save three starving people from dying, it nevertheless is the right thing to do.

		\subsubsection*{Consequences of pathocentric thinking}

		By considering sentiocentrism as our standpoint, it is obvious that all forms of pain should be diminished as much as possible, at least as an utilitiarist. This implies that torturing any animal or keeping them in poor conditions must be disallowed. 

		Some philosophers claim that it is not strictly necessary to make an effort to do good, as long as we don't do anything evil. This is a fallacy, as there is no direct implication from one to the other. If you don't do evil stuff, you're at least not evil, but if you have a moral obligation to help others: say the situation requires it, and you choose not to do anything; then you are not doing good. Kant already showed this difference between moral obligations and good actions. What he didn't consider, though, was the present state of things. This makes a difference, as even if the actions themselves are good, they might be inherently bad due to the circumstance in which they occur.

		\citet{singeryoutube} has reflected on this, and says that animals like cows shouldn't be slaughtered if they are unhappy. Nevertheless, he considers that if a cow lives a happy life, then it doesn't make much of a difference whether they die for our benefit or not. After all, he says, the cow won't feel any pain after being dead, so we could eat a cow that lived happily without any moral rulebreaking.

		While having said that, he also mentions the production of cows being bad for the environment, and, since a higher demand requires a higher supply, the consequence is that eating/buying meat necessarily affects the greenhouse gas emissions. Consequently, he does not support the consumption of any animal on either of these bases.

		As a bonus, from a deontological pathological point of view, it would be considered objectification to produce animals just for our consumption. This means that by refusing to treat them as subjects, we are using them as mere tools for our means. Since this goes against the Universal Moral Law of Nature \autocite{universallaw}, as we wouldn't want to be treated as tools ---much like women or racialised people wouldn't want to be treated as such---, it is morally illegal from a Kantian standpoint too.

 	\printbibliography
\end{document}