---
title: "assignment 2 - mazunki hoksaas"
mainfont: LiberationSerif
date: "2023-09-15"
monofont: LiberationMono
sansfont: LiberationSans
---

## we do a little trolling, tee-hee

first we load the font libraries, since adobe helvetica doesn't seem to be loaded otherwise? kinda weird if you ask me.
```{r include=FALSE}
library(showtext)
showtext_auto()
```

then we load in the csv from a csv file (for some reason loading it as a string as we did last time doesn't work in the notebook?)
```{r}
WIDTH <- 100
HEIGHT <- 100
coords <- read.table("magic.csv", sep = ",", col.names = c("x", "y"));
img <- matrix(255, nrow = HEIGHT, ncol = WIDTH)
for (i in 1:nrow(coords)) {
  img[coords$y[i], coords$x[i]] <- 0
}
```

and we finally plot it, this time with `asp=1` to keep the aspect ratio as the original. when we plotted the image from the interpreter, unlike through Rmarkdown, this wasn't necessary; but is probably good practice to specify anyway
```{r, echo=FALSE}
p <- plot(1, 1, xlim=c(0, WIDTH), ylim=c(0, HEIGHT), type="n", xlab="goof around", ylab="find out", asp=1)
points(coords$x, HEIGHT-coords$y, pch=19, col="black")
```
